DRAFT DRAFT Satisfiability Modulo Theories Competition (SMT-COMP) 2012: Rules and Procedures

### 2012-01-23 rules/rules12.tex
Draft rules and call documents

\date{}

\title{Satisfiability Modulo Theories Competition (SMT-COMP) 2011: Rules and 
Procedures
}

@@

\author{
Roberto Bruttomesso\\
Computer Science\\
USI, Lugano (Switzerland)
\and
Morgan Deters \\
Computer Science \\
New York University \\
\and
Alberto Griggio\\
ES Division\\
FBK, Trento (Italy)
}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR}).
SMT-COMP 2011 is affiliated with the 23nd International
Conference on Computer Aided Verification~(CAV).

The rest of this document, updated from the last year's
version\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, and Aaron Stump.},
describes the rules and competition procedures for SMT-COMP~2011.
The principal change from last year's version is:

\begin{itemize}
\item the introduction of an ``application track,'' based on incremental
      model-checking queries, to promote the integration of SMT solvers
      within actual verification frameworks;
%\item updates to the benchmark selection algorithm, to reduce
%      overwhelming a competition division with a single ``family'' of
%      benchmarks.
% rb 25/3: see later
%\item updates to the difficulty calculation;
%\item a track for parallel solvers, to run before the regular competition.
%\item 5 random ``check'' benchmarks will be generated using Robert
%  Brummayer's SMT fuzzing tool, after all the solvers are uploaded.
%  These will be included in the competition, to help promote attention
%  to the problem of solver correctness (cf.~\cite{brummayer+09}).
\end{itemize}

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted using the SMT-Exec service.  SMT-Exec enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a 12-node computer
cluster purchased with funds from a National Science Foundation
Computing Research Infrastructure grant.  SMT-Exec is provided free of
charge, but it does require a minimal registration, which verifies an
email address and prevents misuse of the service.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to SMT-Exec.  SMT-Exec provides a variety of tabular
and graphical displays of results of solver executions, for
comparison. For the main (sequential-solver)
% rb 23/3: I guess also application track runs on 
% the same clusters as sequential
and application tracks 
of SMT-COMP 2011, SMT-Exec will be configured so that jobs are run on a 64-bit,
uniprocessor Linux kernel.  For the parallel-solver track, SMT-Exec
will be configured so that jobs are run on a 64-bit, multiprocessor
Linux kernel.\footnote{This year, the competition may take
  advantage of off-site computing resources not normally part of
  SMT-Exec; please stay tuned to the SMT-COMP mailing list for final
  details regarding the effective SMT-COMP 2011 competition cluster.}

For participation in SMT-COMP, a solver must be
uploaded as a ``competition'' solver via the usual SMT-Exec upload
mechanism, or, alternatively, a previously-uploaded solver may be
marked as a ``competition'' solver.  In either case, \emph{uploads
must be marked for competition before the deadline;} uploading a solver
to SMT-Exec is not sufficient for competition entry if it's not marked
as being a competition entrant.  The md5 checksums of competition
submissions will be public immediately after the deadline for competition
entry has passed to ensure transparency, and the submissions themselves
will be made public after the competition.  Source code need not be
provided.  However, in order to encourage sharing of source code, extra
recognition will be given to solvers providing source code distributions
including recognition for the top such solver in each division.  See
SMT-Exec~(\url{www.smtexec.org}) for instructions on uploading solvers, as well
as machine specifications.

\header{System description.} %
As part of their submission via SMT-Exec,
SMT-COMP entrants must also include a short (1--2 pages) description of
the system.  This should include a list of all authors of the system
and their present institutional affiliations.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool, but this is
optional.  System descriptions must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

The SMT-Exec upload system will ask for the system description and
the random seed when a solver is marked for competition, so
their inclusion in the uploaded archive itself is optional.

@@

\header{Wrapper tools.} %
A \emph{wrapper tool} is defined as any tool
which calls an SMT solver not written by the author of the wrapper
tool.  The other solver is called the \emph{wrapped tool}.  There are
several rules governing wrapper tools.  For the purposes of these rules,
multiple versions of a wrapped tool are considered different tools.
The goal of these rules is to require wrapper tools to outperform the
tools they wrap (since otherwise, there is no apparent quantitative
way to argue that the wrapper tool improves upon the wrapped tool).

@@

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via SMT-Exec by 7pm, Eastern U.S.
time, July~9th, 2011.  At that time the SMT-Exec service will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until 7pm, Eastern U.S. time, July~11th, 2011.  We strongly encourage
participants to use this two day grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using SMT-Exec after the original deadline on July~9th.

The versions that are present on SMT-Exec at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the morning of July~12th.  See
Section~\ref{sec:timeline} below for a full timeline.

@@

\paragraph{Parallel solver track.} %
Solvers employing concurrency are invited to participate in a
demonstration that will be run before the regular competition.
Parallel solvers must be submitted by the same July~9th deadline,
with the same two-day grace period for resubmission as in the regular
competition.  Submissions to both the parallel solver track
and the main competition are independent: participants must submit
explicitly to both events to participate in both, and they may submit
different (or differently configured) solvers to each.  Benchmarks may
or may not be scrambled for this division.  Entrants should still
include a system description, as for the regular competition.

\medskip
\noindent
Note that the SMT-COMP organizers are considering ways to include
other (``historical'' or otherwise relevant) solvers in all the competition tracks for
demonstration and comparison.  The organizers reserve the right to
include or exclude such solvers, and to make simple modifications
to historical sequential solvers to (na\"ively) take advantage of
multiple processors 
or to connect them with a parser for the SMT-LIB version 2.0 format
(\eg\ by implementing a wrapper tool).

@@

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the course of CAV 2011, from July~14th to~20th.  Results will be
announced in a special session of CAV, on the last day of the
conference, as well as on the SMT-COMP web site.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.
%
If there are enough competitors to run the parallel track, we intend
to run it before or after the main competition (depending on the
estimated duration of the track).

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command.
% rb 25/3: added
% \item A (single) \akey{set-option :print-success false} command, sent just
%       after $\akey{set-logic}$. This is to
%       avoid the output ``\texttt{success}'' to interfere with the
%       answer of the solver to the benchmark. 
\item The $\akey{exit}$ command is the last command.
\item There is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.
\item No other commands besides the ones just mentioned may be used.
\end{enumerate}

% \noindent Note that the final version of the 2.0 format will be frozen
% in early March.  
\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-Exec.
Example benchmark scripts for several Problem Divisions are reported in the Appendix.
Note that they are provided for illustrative purposes only: 
please refer to the SMT-LIB format specification 
and the above definition of benchmark script for the official specification 
of the input format for SMT-COMP.

\medskip
Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', in lowercase, without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.
For more detailed information on the output format, see the
description on the SMT-Exec ``Upload a Solver'' page.

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given formula, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be around 20 minutes.  Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process, using the TreeLimitedRun script, developed for
the CASC competition and available on the SMT-COMP web page.  A
timeout scores the same as if the output is ``\texttt{unknown}''.

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
An abort scores the same as if the output is ``\texttt{unknown}''. 
% rb 25/3: added
Also, as a further measure to prevent misjudgments of solvers,
the first burst of  ``\texttt{success}'' outputs will be 
ignored\footnote{
Note that SMT-LIBv2 requires to produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false: ignoring the first burst of
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of SMT-COMP 2010 to run without changes.}.

@@

The application track aims at evaluating SMT solvers when interacting
with an external verification framework, \eg, a model
checker. This interaction, ideally, happens by means of an online
communication between the model checker and the solver: the model
checker repeatedly sends queries to the SMT solver, which in turn
answers either sat or unsat.  In this interaction an SMT-solver is
required to accept queries incrementally via its standard input channel.

In order to facilitate the evaluation of the solvers in this track, we
propose to set up a ``simulation'' of the aforementioned interaction. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviours of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehaviour, etc. that may happen on the model checker side.
\end{itemize}

@@

\medskip
Solvers must respond immediately to the commands sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification: that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
and with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).

\header{Scoring.}
The score for each benchmark is a pair $\langle n,m\rangle$, with
$n\in[0,N]$ an integral number of points scored for the benchmark,
where $N$ is the number of \akey{check-sat} commands
in the benchmark.  $m\in[0,T]$ is the (real-valued) time is seconds, where $T$ is
the timeout.  The score for the benchmark is initialized with
$\langle0,0\rangle$ and then computed as follows.
\begin{itemize}
\item A correctly-reported \texttt{sat} or \texttt{unsat} answer after
  $s$~seconds (counting from the beginning of this particular
  \akey{check-sat}) contributes $\langle1,s\rangle$ to the running
  score.
\item An answer of \texttt{unknown}, an unexpected answer, a crash, or a memory-out during
  execution of the query, or a benchmark timeout, aborts the execution
  of the benchmark and assigns the current value of the running score
  to the benchmark.  (Recall that there is one timeout for the entire
  benchmark; there are no individual timeouts for queries.)
\item The first incorrect answer to a \akey{check-sat} command has the effect of terminating the
  trace executor, and the returned score for the overall benchmark
  will be $\langle0,0\rangle$, effectively canceling the running score.
\end{itemize}

For example, if a benchmark has 5~\akey{check-sat} commands, and a
timeout of 100~seconds, and a solver solves the first four in
10~seconds each, then times out on the fifth, then the solver's score
is $\langle4,40\rangle$.  If another solver solves each of the first
four in 10~seconds each, and the fifth in another 40~seconds, its
score is $\langle5,80\rangle$.  If a third solver solves the first
four queries in 2~seconds each, but incorrectly answers the fifth, its
score is $\langle0,0\rangle$.

Scores are compared lexicographically---a solver with a higher
$n$-value wins, with the cumulative time only used to break ties.
Benchmarks' scores are summed componentwise to form a solver's total
score for the competition.

@@

We expect the problem divisions for SMT-COMP 2011 to include the following
SMT-LIB \emph{logics}.  These logics are specified in SMT-LIB format on the
SMT-LIB web page.  Note that the ``QF\_'' prefix means the division's
formulas are quantifier-free.  However, the organizers reserve the right to add
(remove) divisions if (not) enough benchmarks and solvers exist for a
particular division.

\begin{itemize}
\item QF\_UF: uninterpreted functions.
\item QF\_RDL: real difference logic.
\item QF\_IDL: integer difference logic.
\item QF\_UFIDL: uninterpreted functions and integer difference logic.
\item QF\_UFLIA: uninterpreted functions and linear integer arithmetic.
\item QF\_UFLRA: uninterpreted functions and linear real arithmetic.
\item QF\_UFNRA: uninterpreted functions and nonlinear real arithmetic.
\item QF\_LRA: linear real arithmetic.
\item QF\_LIA: linear integer arithmetic.
\item QF\_NIA: nonlinear integer arithmetic.
\item QF\_AX: arrays with extensionality.
\item QF\_AUFLIA: arrays, uninterpreted functions and linear integer 
                  arithmetic.
\item QF\_BV: fixed-width bitvectors.
\item QF\_AUFBV: arrays, fixed-width bitvectors and uninterpreted
functions.
\item LRA: (quantified) linear real arithmetic.
\item AUFLIA$+p$: (quantified) arrays, uninterpreted functions and 
linear integer arithmetic, patterns included.
\item AUFLIA$-p$: (quantified) arrays, uninterpreted functions and 
linear integer arithmetic, patterns not included.
\item AUFLIRA: (quantified) arrays, uninterpreted functions and 
mixed linear integer and real arithmetic.
\item UFNIA$+p$: (quantified) uninterpreted functions and 
nonlinear integer arithmetic, patterns included.
\item AUFNIRA: (quantified) arrays, uninterpreted functions and 
mixed nonlinear integer and real arithmetic.
\end{itemize}

@@

\header{Benchmark sources.} %
Benchmark formulas for these divisions
will be drawn from the SMT-LIB library.  Any benchmarks added to
SMT-LIB by the May 15th release (see the timeline in
Section~\ref{sec:timeline}) will be considered eligible.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

>>>>

\date{\small This version revised \the\year-\the\month-\the\day}

\title{DRAFT DRAFT Satisfiability Modulo Theories Competition (SMT-COMP) 2012: Rules and 
Procedures
}

@@

\author{
Roberto Bruttomesso\\
Computer Science\\
USI, Lugano (Switzerland)
\and
David R. Cok \\
GrammaTech, Inc. \\
Ithaca, NY USA \\
\and
Alberto Griggio\\
ES Division\\
FBK, Trento (Italy)
}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR}).
SMT-COMP 2012 is affiliated with the SMT workshop (\url {http://smt2012.loria.fr/}) at the 6th International
Joint Conference on Automated Reasoning~(IJCAR) (\url{http://ijcar.cs.manchester.ac.uk/}).

The rest of this document, updated from the last year's
version\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, and Morgan Deters.},
describes the rules and competition procedures for SMT-COMP~2012.
The principal changes from last year's version are the following:

\begin{itemize}

\item We are concentrating on just a few of the benchmark divisions this year.
  Some divisions, such as QF\_UF, are subsumed into more expressive
  logics; others have received only light interest in past competitions. Our goal is to
  focus the competition on divisions of particular interest to
  applications. To encourage new entrants, we will also accept
  solvers against any benchmark class in an exhibition category:
  results will be displayed and publicly reported.

\item We are retaining and accenting last year's ``application track" and
  encourage submission of benchmarks and competition against these
  benchmarks.

\item The penalty for producing an incorrect result is increased. Competitors
  producing all correct results are ranked above those producing
  incorrect results (a ``soft" disqualification for unsound tools).

\item We are adding an ``unsat core" division of QF\_BV.

\item We will have an exhibition division of solvers that produce proofs.

\item The competition will be run with a new StarExec service, replacing
  the previous SMT-Exec service. At this writing, the new service is
  still being completed and tested; it will be announced later. (The
  previous service will be maintained in reserve.)

\end{itemize}

Additionally, non-competitive divisions and tracks will not be exhibited as
part of the competition: at least two competitors must be entered
into a division for it to run as part of the competition.
If necessary, and based on final submissions, the organizers may elect to
combine competition divisions to make them competitive (such decisions will
be made with input from the community).

As in SMT-COMP 2011, the 2012 version will incorporate a small number of
random ``fuzzer-generated" instances (see below) to help promote
attention to solver correctness.

It is important for competitors to track discussions on the SMT-COMP mailing
list, as clarifications and any updates to these rules will be posted there.

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted using the StarExec\footnote{StarExec is still under development as of 21 January 2012;
updates on its progress and use will be posted on www.smtcomp.org and through
the SMT-COMP mailing list. Statements in these rules may need correction once the UI for the service is completed.} service.  
StarExec enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a computer
cluster purchased with funds from a National Science Foundation
Computing Research Infrastructure grant.  StarExec is provided free of
charge, but it does require a minimal registration, which verifies an
email address and prevents misuse of the service.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to StarExec.  StarExec provides a variety of tabular
and graphical displays of results of solver executions, for
comparison. For the main (sequential-solver)
% rb 23/3: I guess also application track runs on 
% the same clusters as sequential
and application tracks 
of SMT-COMP 2012, StarExec will be configured so that jobs are run on a 64-bit,
uniprocessor Linux kernel.  For the parallel-solver track, StarExec
will be configured so that jobs are run on a 64-bit, multiprocessor
Linux kernel.

For participation in SMT-COMP, a solver must be
uploaded as a ``competition'' solver via the StarExec upload
mechanism, or, alternatively, a previously-uploaded solver may be
marked as a ``competition'' solver.  In either case, \emph{uploads
must be marked for competition before the deadline;} uploading a solver
to StarExec is not sufficient for competition entry if it's not marked
as being a competition entrant.  The md5 checksums of competition
submissions will be public immediately after the deadline for competition
entry has passed to ensure transparency, and the submissions themselves
will be made public after the competition.  Source code need not be
provided.  However, in order to encourage sharing of source code, extra
recognition will be given to solvers providing source code distributions
including recognition for the top such solver in each division.  Instructions
for uploading solvers and machine specifications will be posted as soon
as they are available.

\header{System description.} %
As part of their submission via StarExec,
SMT-COMP entrants must also include a short (1--2 pages) description of
the system.  This should include a list of all authors of the system
and their present institutional affiliations.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool, but this is
optional.  System descriptions must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

The StarExec upload system will ask for the system description and
the random seed when a solver is marked for competition, so
their inclusion in the uploaded archive itself is optional.

@@

\header{Wrapper tools.} %
A \emph{wrapper tool} is defined as any tool
which calls an SMT solver not written by the author of the wrapper
tool.  The other solver is called the \emph{wrapped tool}.  There are
several rules governing wrapper tools that wrap tools that have been or could be submitted as independent entrants.\footnote{A wrapper
for a wrapped tool that does not accept SMT-LIBv2 format is considered to simply be an SMT version of the underlying tool.}
For the purposes of these rules,
multiple versions of a wrapped tool are considered different tools.
The goal of these rules is to require wrapper tools to outperform the
tools they wrap (since otherwise, there is no apparent quantitative
way to argue that the wrapper tool improves upon the wrapped tool).

@@

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec by 7pm, Eastern U.S.
time, June~15, 2012.  At that time the StarExec service will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until 7pm, Eastern U.S. time, June~18, 2012.  We strongly encourage
participants to use this weekend grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the original deadline on June~15.

The versions that are present on StarExec at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the morning of June~22.  See
Section~\ref{sec:timeline} below for a full timeline.

@@

\paragraph{Parallel solver track.} %
Solvers employing concurrency are invited to participate in a
demonstration that will be run before the regular competition.
Parallel solvers must be submitted by the same June~15th deadline,
with the same grace period for resubmission as in the regular
competition.  Submissions to both the parallel solver track
and the main competition are independent: participants must submit
explicitly to both events to participate in both, and they may submit
different (or differently configured) solvers to each.  Benchmarks may
or may not be scrambled for this division.  Entrants should still
include a system description, as for the regular competition.

\medskip
\noindent
Note that the SMT-COMP organizers may include
other (``historical'' or otherwise relevant) solvers in all the competition tracks for
demonstration and comparison.  The organizers reserve the right to
include or exclude such solvers, and to make simple modifications
to historical sequential solvers to (na\"ively) take advantage of
multiple processors 
or to connect them with a parser for the SMT-LIB version 2.0 format
(\eg\ by implementing a wrapper tool).

@@

Solvers will be publicly evaluated in the following tracks, listed here and 
described in detail below. In exhibition tracks, solvers are evaluated and results posted, but no awards are announced. The other tracks are competition tracks; if there are sufficient entrants, a winner will be announced.
\begin{itemize}
\item a main track: sequential execution evaluated separately on benchmarks from each of several different logics
\item a parallel track: parallel execution evaluated on the same benchmarks and the same divisions as the main track
\item application track: evaluation on command scripts [TODO: FOR WHICH LOGICS]
\item exhibition track: solvers (particularly new entrants) may be submitted against any benchmark division; the solver will be evaluated and the results posted, though not included in competition
\item unsat-core track: a competition among solvers capable of computing unsat cores
\item proof-generation exhibition: an exhibition track for solvers capable of generating proofs
\end{itemize}

\subsection{Logistics}

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the course of IJCAR 2012, from June~26 to the 29th.  Results will be
announced in a special session of IJCAR, on the last day of the
conference, as well as at the SMT workshop and on the SMT-COMP web site.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

If there are enough competitors to run the parallel track, we intend
to run it before or after the main competition (depending on the
estimated duration of the track).

The organizers will prioritize the running of the competition tracks, and may shift the
time period or order of the competition or exhibition tracks in order to complete SMT-COMP
in the course of the IJCAR conference.

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command.
% rb 25/3: added
% \item A (single) \akey{set-option :print-success false} command, sent just
%       after $\akey{set-logic}$. This is to
%       avoid the output ``\texttt{success}'' to interfere with the
%       answer of the solver to the benchmark. 
\item The $\akey{exit}$ command is the last command.
\item There is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item There is at most one $\akey{set-info}$ command for \texttt{status}.
%%%%%%%%%%%%%%%%%%% TODO: Need to define the set-info command.
\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item Extra symbols are declared exactly once before any
  use, using $\akey{declare-sort}$
  %, $\akey{define-sort}$,
  %$\akey{declare-fun}$, or $\akey{define-fun}$.  
  or $\akey{declare-fun}$.
  They must be part of the allowed signature expansion for the logic.
  Moreover, all sorts declared with a $\akey{declare-sort}$ command must have zero arity.
\item In the unsat-core competition, the $\akey{set-logic}$ command is preceded by a 
$\akey{set-option :produce-unsat-cores true}$ command and the 
$\akey{check-sat}$ command is followed by a $\akey{get-unsat-core}$ command.
Also, some or all of the $\akey{assert}$ commands will assert named formula
(for example, \akey{assert (! P :named F1)}). The $\akey{get-unsat-core}$
command must return a parenthesized list of formula names, as specified by the 
SMTLIBv2 standard.
\item In the proof generation competition, the $\akey{set-logic}$ command is preceded by a 
$\akey{set-option :produce-proofs true}$ command and the 
$\akey{check-sat}$ command is followed by a $\akey{get-proof}$ command.
The $\akey{get-proof}$ command must return a proof in a solver-dependent format as an
S-expression.
\item No other commands besides the ones just mentioned may be used.

\end{enumerate}

\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-Exec.
Example benchmark scripts for several Problem Divisions are reported in the Appendix.
Note that they are provided for illustrative purposes only: 
please refer to the SMT-LIB format specification 
and the above definition of benchmark script for the official specification 
of the input format for SMT-COMP.

@@

The main track competition will consist of selected benchmarks in each of the
logic divisions given below.
Each benchmark script will be presented to the standard input of the solver.
Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', in lowercase, without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.
For more detailed information on the output format, see the
description on the StarExec ``Upload a Solver'' page.  %%%%%%% TODO: CHeck this and all referecnes to StarExec

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given formula, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 20 minutes, as it was in 2011.  Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process, using the TreeLimitedRun script, developed for
the CASC competition and available on the SMT-COMP web page.  A
timeout scores the same as if the output is ``\texttt{unknown}''.
%%%%%%%%%%% TODO: Clarify time limit

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
An abort scores the same as if the output is ``\texttt{unknown}''. 
% rb 25/3: added
Also, as a further measure to prevent misjudgments of solvers,
any  ``\texttt{success}'' outputs will be 
ignored\footnote{
Note that SMT-LIBv2 requires to produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false; ignoring the
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of previous competitions to run without changes.}.

@@

The application track evaluates SMT solvers when interacting
with an external verification framework, \eg, a model
checker. This interaction, ideally, happens by means of an online
communication between the model checker and the solver: the model
checker repeatedly sends queries to the SMT solver, which in turn
answers either \texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its standard input channel.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction, as was done in 2011. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviours of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehaviour, etc. that may happen on the model checker side.
\end{itemize}

@@

\medskip
Solvers must respond immediately to the commands sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
and with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).  We anticipate the
timeout to be around 30~minutes (as it was in 2011).

\subsection{Exhibition track}

To encourage new entrants in any logic, we will have an exhibition track.
Any solver submitted for ``competition'' in a non-competitive track will
be executed on benchmarks for the logic divisions chosen for that solver.
The results will be made public and the solver acknowledged. Any logic divisions
for which there are benchmarks may be specified.

\subsection{Unsat-core and proof-generation tracks}

Applications such as software verification and model checking are enhanced by having
proofs and unsatisfiable cores available from SMT solvers. Accordingly we are 
encouraging solvers to add such capabilities by recognizing them in SMTCOMP 2012.

The unsat-core track will evaluate solvers' capability to generate unsatisfiable cores for
problems that are known to be unsatisfiable. Solvers will be measured by the smallness of
the unsat core they return. The SMT-LIBv2 language accommodates this functionality by 
providing two features: the ability to name top-level (asserted) formula and the ability to
request an unsat-core after a check-sat command returns \texttt{unsat}. The unsat-core that
is returned consists of a list of names of formula. In the competition we will check that the
returned unsat-core is well-defined and is still unsatisfiable.

Similarly, the competition will feature an exhibition track of proof-generating solvers. This is
simply an exhibition track because there is as yet no SMT-LIB standard way to express proofs. We hope
that encouraging the capability in solvers will also encourage a common proof format. The exhibition
will count the number of benchmarks for which a solver successfully generates a proof; the proofs 
themselves may be spot checked by hand or in a semi-automated fashion, depending on the format.

TODO: What benchmark logics will be used.

@@

We expect the problem divisions for SMT-COMP 2012 to include the following
SMT-LIB \emph{logics}.  These logics are specified in SMT-LIB format on the
SMT-LIB web page.  Note that the ``QF\_'' prefix means the division's
formulas are quantifier-free.  However, the organizers reserve the right to add
(remove) divisions if (not) enough benchmarks and solvers exist for a
particular division.

Note that this year we are reducing the number of competitive benchmark divisions.
Our reasons are to concentrate the competition and to focus development and attention
on benchmarks that are more challenging and relevant to applications.

\begin{itemize}

\item QF\_BV: fixed-width bitvectors. This logic is important to ``bit-blasting'' model checking of software.
\item QF\_AUFBV: arrays, fixed-width bitvectors and uninterpreted
functions. This logic is also key to software model-checking, but adds arrays (which can be used for memory models) and uninterpreted functions.

\item QF\_UFLIA: uninterpreted functions and linear integer arithmetic (benchmarks may also be taken from  QF\_LIA). This division evaluates reasoning about integers.
\item QF\_UFLRA: uninterpreted functions and linear real arithmetic (benchmarks may also be taken from  QF\_LRA). This division evaluates reasoning about real numbers.

\item AUFLIA$+p$: (quantified) arrays, uninterpreted functions and 
linear integer arithmetic, patterns included. This and the subsequent divisions evaluate solvers' abilities to handle quantified formula.
\item AUFLIA$-p$: (quantified) arrays, uninterpreted functions and 
linear integer arithmetic, patterns not included.
\item AUFNIRA: (quantified) arrays, uninterpreted functions and 
mixed nonlinear integer and real arithmetic

\end{itemize}

@@

\header{Benchmark sources.} %
Benchmark formulas for these divisions
will be drawn from the SMT-LIB library.  Any benchmarks added to
SMT-LIB by the April 15th release (see the timeline in
Section~\ref{sec:timeline}) will be considered eligible.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.
